A Highly Parallel Approach for Solving Computationally Expensive Multicriteria Optimization Problems

### 2019-11-19 2019/RSCD_Gergel_Kozinov/RSCD_2019.tex
Добавил исходники статей по ГО за 2019г. (Гергель Козинов)

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%%!PS-Adobe-3.0 EPSF-3.0
%%%BoundingBox: 19 19 221 221
%%%CreationDate: Mon Sep 29 1997
%%%Creator: programmed by hand (JK)
%%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[runningheads]{llncs}
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
%
\usepackage{graphicx}
\usepackage{mathtools,amssymb}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{A Highly Parallel Approach for Solving Computationally Expensive Multicriteria Optimization Problems
\thanks{This work has been supported by Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using superior-performance supercomputers.''}
}

@@

\author{Victor Gergel\orcidID{0000-0002-4013-2329} \and Evgeny Kozinov \orcidID{0000-0001-6776-0096}}

@@

A general approach to solving the MCO problem used in this paper consists in the replacement of solving the original MCO problems by solving a series of simpler optimization problems: 
\begin{equation}\label{eq:3}
\begin{split}
\min{(\phi(x) = F(\lambda, y(x)), y(x) \in D, x \in [0,1])},\\
F(\lambda,y(x))=\max⁡{(\lambda_i f_i(y(x)), 1 \leq i \leq s)},\\
\lambda=(\lambda_1,\lambda_2,\dots,\lambda_s) \in \Lambda \subset R^s : \sum_{i=1}^s{\lambda_i}=1, \lambda_i \geq 0, 1 \leq i \leq s
\end{split}
\end{equation}
where $F(\lambda,y(x))$ is the minimax convolution of the efficiency criteria of the MCO problem \cite{c2,c6,c19} with the use of the vector of the convolution coefficients $\lambda \in \Lambda$ and $y(x)$ is a continuous and unambiguous mapping of the interval [0,1] onto an $N$-dimensional search domain $D$ -- see, for example, \cite{c7,c8}.

@@

Multiextremal optimization is a research area being actively developed -- its state of the art is presented, for example, in \cite{c7,c9,c10}. The information-statistical theory of global optimization is considered to be one of the promising approaches \cite{c7}. HPC systems are used widely for solving time-consuming global search problems \cite{c7,c11,c14,c15}.

@@

Within the proposed approach, for solving the reduced one-dimensional multiextremal optimization subproblems (\ref{eq:3}) it is proposed to use the well-known Parallel Algorithm of Global Search (PAGS) developed within the framework of the information-statistical theory of the multiextremal optimization \cite{c7}. This method has a good theoretical background and has demonstrated its efficiency as compared to other global search algorithms (see also the results of numerical experiments in Section \ref{sec:4}).

@@

Rule 5. Perform new trials at the points $x^{k+j}$, $1 \leq j \leq p$ placed in the intervals with the maximum characteristics from (\ref{eq:11}) according to the expressions

@@

As the current estimate of the optimization problem solution, the minimum computed value of the optimization function is accepted i. e.
\begin{eqnarray} \label{eq:14}
z_k^*=\min⁡{\{z_i:1\leq i \leq k\}}.
\end{eqnarray}

Within the framework of the proposed approach, PAGS is applied to solving the MCO problem (\ref{eq:1}) according to the scheme (\ref{eq:3}). The availability of SIS from (\ref{eq:6}) allows transforming the results of the previous computations to the values of a next optimization problem (\ref{eq:3}) without any time-consuming computations of the efficiency criteria values for new values of the convolution coefficients, i.e.
\begin{eqnarray} \label{eq:15}
z_i'=\max{⁡(\lambda_j' f_i^j,1\leq j \leq s)},1 \leq i \leq k.
\end{eqnarray}

@@

To solve each subproblems $F(\lambda,y(x))$ in (\ref{eq:3}), the termination condition (\ref{eq:13}) was replaced by the termination condition by the condition
\begin{equation*}
F(\lambda,y(x^{k+j})) \leq F(\lambda,y^* )+\varepsilon,1 \leq j \leq p,  F(\lambda,y^*) = \min_{y\in D_\delta}⁡{F(\lambda,y)}.
\end{equation*}

After solving subproblems (\ref{eq:3}) for 50 values of $\lambda$ a numerical approximation of the Pareto domain $PD_{\Omega_k}(f,D)$ is constructed using the search information accumulated in $\Omega_k$ from (\ref{eq:6}) as result of the solution of the MCO problem. The resulting estimation of the accuracy of the calculated approximation $PD_{\Omega_k}(f,D)$ is defined as the closeness to the $\delta$-approximation $PD_\delta(f,D)$ of the Pareto domain of the MCO problem 
\begin{equation}\label{eq:18}
\Delta=\frac{1}{M} \sum_{y\in PD_\delta(f,D)}{d(y)}, d(y)=\min_{\overline{y}\in PD_{\Omega_k}(f,D)}{⁡\|y-\overline{y}\|},
\end{equation}
where $\|.\|$ denotes the norm in the $R^N$ space, and $M$ is the number of points used in the approximation $PD_{\delta}(f,D)$.

@@

@@

\bibitem{c15}	Gergel, V.: An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems. In: 2nd International Conference on Mathematics and Com-puters in Sciences and in Industry, pp. 13--18 (2015)

@@

\bibitem{c50}	Borisenko, A., Gorlatch, S.: Parallelizing Metaheuristics for Optimal Design of Multiproduct Batch Plants on GPU. In: LNCS, Springer, Cham, \textbf{10421}, pp. 405--417 (2017)

>>>>

\documentclass[runningheads]{llncs}
%
%
\usepackage{graphicx}
\usepackage{mathtools,amssymb}
\usepackage{bbding}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{A Highly Parallel Approach for Solving Computationally Expensive Multicriteria Optimization Problems}

@@

\author{Victor Gergel\orcidID{0000-0002-4013-2329} (\Envelope) \and Evgeny Kozinov \orcidID{0000-0001-6776-0096}}

@@

A general approach to solving the MCO problem used in this paper consists in the replacement of solving the original MCO problems by solving a series of simpler optimization problems: 
\begin{equation}\label{eq:3}
\begin{split}
\min{(\phi(x) = F(\lambda, y(x)), y(x) \in D, x \in [0,1])},\\
F(\lambda,y(x))=\max{(\lambda_i f_i(y(x)), 1 \leq i \leq s)},\\
\lambda=(\lambda_1,\lambda_2,\dots,\lambda_s) \in \Lambda \subset R^s : \sum_{i=1}^s{\lambda_i}=1, \lambda_i \geq 0, 1 \leq i \leq s
\end{split}
\end{equation}
where $F(\lambda,y(x))$ is the minimax convolution of the efficiency criteria of the MCO problem \cite{c2,c6,c19} with the use of the vector of the convolution coefficients $\lambda \in \Lambda$ and $y(x)$ is a continuous and unambiguous mapping of the interval [0,1] onto an $N$-dimensional search domain $D$ -- see, for example, \cite{c7,c8}.

@@

Multiextremal optimization is a research area being actively developed -- its state of the art is presented, for example, in \cite{c7,c9,c10,c60}. The information-statistical theory of global optimization is considered to be one of the promising approaches \cite{c7}. HPC systems are used widely for solving time-consuming global search problems \cite{c7,c11,c14,c15}.

@@

Within the proposed approach, for solving the reduced one-dimensional multiextremal optimization subproblems (\ref{eq:3}) it is proposed to use the well-known Parallel Algorithm of Global Search (PAGS) developed within the framework of the information-statistical theory of the multiextremal optimization \cite{c7,c25}. This method has a good theoretical background and has demonstrated its efficiency as compared to other global search algorithms (see also the results of numerical experiments in Section \ref{sec:4}).

@@

\textit{Rule 5.} Perform new trials at the points $x^{k+j}$, $1 \leq j \leq p$ placed in the intervals with the maximum characteristics from (\ref{eq:11}) according to the expressions

@@

As the current estimate of the optimization problem solution, the minimum computed value of the optimization function is accepted i. e.
\begin{eqnarray} \label{eq:14}
z_k^*=\min{\{z_i:1\leq i \leq k\}}.
\end{eqnarray}

Within the framework of the proposed approach, PAGS is applied to solving the MCO problem (\ref{eq:1}) according to the scheme (\ref{eq:3}). The availability of SIS from (\ref{eq:6}) allows transforming the results of the previous computations to the values of a next optimization problem (\ref{eq:3}) without any time-consuming computations of the efficiency criteria values for new values of the convolution coefficients, i.e.
\begin{eqnarray} \label{eq:15}
z_i'=\max{(\lambda_j' f_i^j,1\leq j \leq s)},1 \leq i \leq k.
\end{eqnarray}

@@

To solve each subproblems $F(\lambda,y(x))$ in (\ref{eq:3}), the termination condition (\ref{eq:13}) was replaced by the termination condition by the condition
\begin{equation*}
F(\lambda,y(x^{k+j})) \leq F(\lambda,y^* )+\varepsilon,1 \leq j \leq p,  F(\lambda,y^*) = \min_{y\in D_\delta}{F(\lambda,y)}.
\end{equation*}

After solving subproblems (\ref{eq:3}) for 50 values of $\lambda$ a numerical approximation of the Pareto domain $PD_{\Omega_k}(f,D)$ is constructed using the search information accumulated in $\Omega_k$ from (\ref{eq:6}) as result of the solution of the MCO problem. The resulting estimation of the accuracy of the calculated approximation $PD_{\Omega_k}(f,D)$ is defined as the closeness to the $\delta$-approximation $PD_\delta(f,D)$ of the Pareto domain of the MCO problem 
\begin{equation}\label{eq:18}
\Delta=\frac{1}{M} \sum_{y\in PD_\delta(f,D)}{d(y)}, d(y)=\min_{\overline{y}\in PD_{\Omega_k}(f,D)}{\|y-\overline{y}\|},
\end{equation}
where $\|.\|$ denotes the norm in the $R^N$ space, and $M$ is the number of points used in the approximation $PD_{\delta}(f,D)$.

@@

\section*{Acknowledgements}
This work has been supported by Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using superior-performance supercomputers.''

@@

\bibitem{c15}	Barkalov, K., Gergel, V., Lebedev, I.: Solving global optimization problems on GPU cluster. AIP Conference Proceedings, \textbf{1738}, pp. 400006 (2016) \doi{10.1063/1.4952194}

@@

\bibitem{c50}	Borisenko, A., Gorlatch, S.: Parallelizing Metaheuristics for Optimal Design of Multiproduct Batch Plants on GPU. In: LNCS, Springer, Cham, \textbf{10421}, pp. 405--417 (2017)
\bibitem{c25}	Strongin, R.G., Gergel, V.P.: Parallel computing for globally optimal decision making. In: Lecture Notes in Computer Science, \textbf{2763}, pp. 76--88 (2003)  
\bibitem{c60}	Sakharov, M.K., Karpenko, A.P.:Adaptive Load Balancing in the Modified Mind Evolutionary Computation Algorithm In: Supercomputing Frontiers and Innovations, \textbf{5}(4), pp. 5--14 (2018)
